Best Practices in Latent Class Analysis using Free Open Source Software

### 2022-12-04 manuscript.Rmd
update 2022-12-04 10:03:24

Another way to conceptualize latent class analysis is by analogy to a measurement model in structural equation modeling.
A mixture model is like confirmatory factor analysis,
except that the continuous latent variable is substituted with a categorical latent variable.
One difference between
the two techniques is that factor analysis can be considered as a way to
group observed *variables* into latent constructs, and LCA groups
*individuals* into classes. While factor analysis seeks to describe the
relations between variables, LCA seeks to describe each unobserved group
and the differences between them. In line with this distinction, LCA
(mixture models) is referred to as a "person-centered" technique, and
factor analysis as a "variable-centered" technique
[@nylund-gibson_ten_2018; @masyn_latent_2013].

@@

Another use case is to explore potential heterogeneity in a population,
and to determine whether it is comprised of latent subpopulations.
For example, researchers suspected the existence of different risk profiles of substance use, sexual behavior, and mental health outcomes, and used LCA to determine how many such risk profiles could be distinguished [@hopfer_social_2014].
Another explored heterogeneity in the population of military service members and veterans with prior suicide attempts,
and identified two latent profiles characterized by high versus low suicide risk based on self-reported suicidal ideation and information about prior attempted methods and severity
[@gromatsky_characteristics_2022].

@@

### Controlling for covariates

An extension of LCA is that containing covariates which can be used to
predict class membership. In this approach, we not only model the latent
class variable based on indicators, but we also relate the class
membership to other explanatory variables (@vermunt_latent_2017). An
example of using covariates comes from @nozadi_moderating_2016 who
applied LCA to identify the probability of children's membership to an
anxiety class. The authors tested several covariates including
children's age, sex, and accuracy scores. Age and sex were not found to
be related to the children's latent class membership, hence these
covariates were excluded from the analysis. Accuracy scores, however,
were related to probabilities of being in anxiety and attention-anxiety
classes. Therefore this covariate was kept as a valuable predictor.
Another example comes from @van_lissa_divergence_2015 who included sex 
as a predictor of the developmental trajectories and class membership in 
the previously discussed adolescent empathy development study. The rationale
was that adding sex as a covariate would account for known sex differences 
in empathy development, and lead to a model with lower bias.

### Distal outcomes

When our interest is the prediction of one or more outcomes, LCA can be used 
to construct latent classes as categorical predictors. @lanza_latent_2013 
demonstrated how LCA can be used to classify adolescents into depression 
classes, and subsequently these classes can be used to predict smoking, grades,
and delinquency. The study showed that the outcomes predicted by class 
membership can be binary (regular smoking), continuous (grades) or 
count (delinquency). In the previously discussed study of Belgian adolescent
identities, @maene_perceived_2022 first constructed the latent variable
representing different multiple identification strategies. Subsequently, 
the most likely class membership based on posterior class probability was
used as an "observed" categorical variable for prediction of depressive symptoms.

@@

@@

The best practices in estimation, as outlined in Table
\@ref(tab:checkest), are rooted in existing recommendations for best
practices for estimating specific subtypes of LCA, including latent
class growth analysis [@van_de_schoot_grolts-checklist_2017] and latent
class analysis with ordinal indicators [e.g., @nylund-gibson_ten_2018].
These were generalized to be relevant to all types of LCA, and updated
to current best practices, as explained below.

@@

Examining observed data is essential for any analysis as it may reveal
patterns and violations of assumptions that had not been considered
prior to data collection. Special attention should be paid to level of
measurement of the indicators. Finite Gaussian mixture models (including
LPA) are only suitable for continuous variables. Indicators with an
ordinal level of measurement are likely to violate the assumption of
within-class normal distributions of mixture models [see
@vermunt_k-means_2011]. Personal experience consulting on LCA methods
and moderating the `tidyLPA` Google group suggest that the
misapplication of mixture models to ordinal (e.g., Likert-type)
indicators is the most common source of user error. Whereas it has been
argued that some parametric methods are robust when scales with 7+
indicators are treated as continuous [e.g., @norman_likert_2010],
this certainly does not imply that all methods are.
It is very unlikely
that an ordinal variable can be modeled as a *mixture* of multiple
normal distributions. The problem becomes egregious when the number of
classes estimated equals or exceeds the number of categories; in this
case, each class-specific mean could describe a single response
category, and a class-specific variance component would be nonsensical.
In sum, Likert-type scales are rarely suitable for Gaussian mixture modeling;
latent class analysis with ordinal indicators is more appropriate.

@@

Previous work has emphasized the importance of examining the pattern of
missing data and reporting how missingness was handled
[@van_de_schoot_grolts-checklist_2017]. Three types of missingness have
been distinguished in the literature [@rubin_inference_1976; 
@enders_applied_2022]: Missing completely at random (MCAR), which means that 
missingness is random and unrelated to any observed or unobserved variables;
missing at random (MAR), which means that missingness is contingent on
the *observed* data (and can thus be accounted for); and finally missing
not at random (MNAR), which means that missingness is related to
unobserved factors. It is possible to conduct a so-called "MCAR" test,
for example the non-parametric MCAR test [@jamshidian_tests_2010]. But
note that the name "MCAR test" is somewhat misleading, as the
null-hypothesis of this test is that the data are not MAR, and a
significant test statistic indicates that missingness is related to the
observed data (MAR). A non-significant test statistic does not
distinguish between MCAR or MNAR. As Little's classic MCAR test relies
on the comparison of variances across groups with different patterns of
missing data, it assumes normality [@little_test_1988]. This assumption
is tenuous in the context of LCA. A non-parametric MCAR test, as
provided by Jamshidian and Jalal, may be more suitable
[@jamshidian_tests_2010]. Unfortunately, this test was removed from the
central R-repository CRAN due to lack of maintenance. For this tutorial,
I have re-implemented it in the `mice` package as `mice::mcar()`, with a
fast backend in C++ and new printing and plotting methods.

While we concur that investigating missingness is due diligence, it is
important to emphasize that missingness is adequately handled by default
in many software packages for LCA, such as Mplus, Latet GOLD, and `OpenMx` 
which is the backend of `tidySEM`. These packages use Full Information Maximum
Likelihood (FIML) estimation, which makes use of all available
information without imputing missing values. FIML is a best-practice
solution for handling missing data; on par with multiple imputation when the 
estimation model includes the same variables as the imputation model
[@lee_comparison_2021]. FIML estimation assumes that missingness is
either MCAR or MAR. Thus, one would typically proceed with FIML,
regardless of the outcome of an MCAR test. Although FIML does not, by
default, handle missingness in exogenous variables -- all indicator
variables in LCA are endogenous, so this is not a concern.
<!-- are you sure Mplus accounts for missing data automatically when the indicators are categorical? I know that is not the case with CFA/SEM models -->
<!-- CJ: yes, it does -->

Multiple imputation is less suitable to LCA for two reasons. First,
because LCA methods are often computationally expensive, and conducting
them on multiple imputed datasets may be unfeasible. Second, because
there is no straightforward way to integrate LCA results across multiple
datasets. To conclude; our recommendation is to inspect missingness
(e.g., using `mice::MCAR()`) and report the proportion of missingness
per variable (e.g., using `tidySEM::descriptives()`), before proceeding
with FIML. One minor concern is that the K-means algorithm, which
`tidySEM` uses for determining starting values, is *not* robust to
missing values. When it fails, `tidySEM` automatically switches to
hierarchical clustering, unless the user specifies a different
clustering algorithm or uses manual starting values.

### Alternative Model Specifications

In order to aid researchers working with latent trajectory models,
@van_de_schoot_grolts-checklist_2017 developed a protocol called
Guidelines for Reporting on Latent Trajectory Studies (GRoLTS). Two of
the GRoLTS guidelines (namely, 6a and 6b) refer to considering
alternative model specifications. Both are discussing specific cases in
latent trajectory model specification. The first is about whether the
variance of the growth parameter is estimated freely or fixed, and the
second is about whether conditional independence is assumed, as the
researcher might also want to free-up the variance-covariance structure.
In LCA, there are also many different ways to specify the model. Means,
variances, and covariances between the indicators can either be
constrained across classes, or estimated freely. Researchers using LCA
should transparently report their chosen parametrization as well as
discuss different parametrizations that were tested in the process.

Different types of latent class models have different parameters. For
example, mixture models and latent profile analyses typically have
class-specific means, variances, and covariances. Latent growth analyses
have the same parameters, but with respect to the latent growth
variables. Latent class analyses with ordinal indicators have
thresholds. All of these parameters can be freely estimated, constrained
to be equal across classes, or fixed to a certain value (e.g., to zero).
The total number of parameters thus scales with the number of estimated
classes. Consequently, LCA methods have a potentially very high number
of parameters. As any of these parameters could be misspecified, it is
important to consider alternative model specifications. However,
alternative model specifications may be approached differently depending
on whether an analysis is data driven (exploratory), or theory driven
(confirmatory). This distinction has remained underemphasized in prior
writing.

Prior literature on LCA has emphasized its exploratory applications
[@nylund_deciding_2007]. In exploratory LCA, a large number of models
are typically estimated in batch, with varying numbers of classes and
model specifications. The "correct" model specification is then
determined based on a combination of fit indices, significance tests,
and interpretability. For latent profile analysis, the function
`tidySEM::mx_profiles(classes, variances, covariances)` largely
automates this process. The argument `classes` indicates which class
solutions should be estimated (e.g., 1 through 6). The argument
`variances` specifies whether variances should be `"equal"` or
`"varying"` across classes. The argument `covariances` specifies whether
covariances should be constrained to `"zero"`, `"equal"` or `"varying"`
across classes. The means are free to vary across classes by default,
although the more general function `tidySEM::mx_mixture()` could be used
to circumvent this. After all models have been estimated, the function
`tidySEM::table_fit()` can be used to obtain a model fit table suitable
for determining the optimal model according to best practices. Note,
however, that this table does not include the bootstrapped likelihood
ratio test (BLRT) by default, because this test is very computationally
expensive. It is recommended to use the function `tidySEM::BLRT()` to
compare a shortlist of likely candidate models based on other fit
indices. More on fit indices can be found in the Model Fit Indices
subsection of this paper.

Confirmatory LCA typically requires less comprehensive alternative model
specifications. For example, in the context of a preregistered analysis,
the main models of interest would have been specified a priori. Even in
this case, the theoretical model could be compared to a few others to
contextualize it.

@@

Among studies using LCA, reporting practices vary significantly
[@weller_latent_2020]. Various authors have tried to improve and
standardize ways of reporting LCA [e.g. @masyn_latent_2013;
@weller_latent_2020], but more work is needed. @van_lissa_worcs_2020
developed WORCS, a workflow for open reproducible code in science. WORCS
consists of step-by-step guidelines for research projects based on the
TOP-guidelines developed by @nosek_promoting_2015. WORCS workflow can be
easily implemented in R in form of an R package which facilitates
preregistration, article drafting, version control, citation and
formatting, among others [@van_lissa_worcs_2020].

TOP-guidelines emphasize the use of comprehensive citation (including
referencing the software used in the analysis), as well as code and data
sharing wherever possible [@nosek_promoting_2015]. @van_lissa_worcs_2020
suggest sharing synthetic data in case the original data cannot be
shared, and provide functions to generate such synthetic data. Ideally,
the entire research project is made reproducible so that others may
download it and reproduce it with just one click; for guidance, see
@van_lissa_worcs_2020.

@@

```{r, include = TRUE, eval=F}
library(tidySEM) # loading tidySEM
df <- data_mix_ordinal  # dataset
df[1:4] <- lapply(df, ordered) # indicators as ordered factors
```

@@

```{r, include = TRUE, eval=F}
tidySEM::descriptives(df) # descriptives
```

@@

Before we fit a series of LCA models, we set a random seed using
`set.seed()`. The seed number can be any digit. This is an important
step as there is some inherent randomness in the LCA computations, and
having the same seed number ensures that two separate researcher obtain
exactly the same results when fitting LCA models.

Finally, we reach the step of fitting LCA models. To do so, we use
`tidySEM::mx_lca()` which takes data and number of classes as inputs. 
Notice that the analysis will be done on all the variable in the data set, 
so you have to make sure it only includes the relevant variables. 
In our example, we want to fit 1 to 4 class solutions and compare 
their output. Depending on your computer's computational power, 
this might take a while.

@@

Notice that the model run 10 times for each $K$ class solution. 
This way we can choose the best fitting model across all the solutions. 
In this example, we see that all models converged without issues.

We can inspect the class of resulting objects.

```{r, include = TRUE, eval=F}
is(res)
is(res[[1]])
```

As `tidySEM::mx_lca()` run several models at a time, the resulting object `res` 
is a `mixture_list` class where each $K$ class solution is a list element, or
more specifically an `OpenMx` model (`MxModel`).

@@

Description <- c(
    "the $K$-class solution",
    "the -2*log-likelihood of each model",
    "Number of estimated parameters in each model",
    "the Akaike Information Criterion value",
    "the Bayesian Information Criterion value",
    "the lowest posterior class probability by most likely class membership",
    "the highest posterior class probability by most likely class membership",
    "the lowest class proportion based on the posterior class probabilities",
    "the highest class proportion based on the posterior class probabilities")

>>>>

Another way to conceptualize latent class analysis is by analogy to a measurement model in structural equation modeling [@molenaarArbitraryNatureLatent1994].
A mixture model is like confirmatory factor analysis,
except that the continuous latent variable is substituted with a categorical latent variable.
One difference between
the two techniques is that factor analysis can be considered as a way to
group observed *variables* into latent constructs, and LCA groups
*individuals* into classes.
While factor analysis seeks to describe the relations between variables,
LCA seeks to describe unobserved groups and classify individual observations to those groups.
In line with this distinction, LCA
(mixture models) is referred to as a "person-centered" technique, and
factor analysis as a "variable-centered" technique
[@nylund-gibson_ten_2018; @masyn_latent_2013].

@@

Another use case is to explore potential heterogeneity in a population,
and to determine whether the population is comprised of latent subpopulations, and if so, how many.
The primary focus in this use case is on class enumeration.
For example, researchers suspected the existence of different risk profiles of substance use, sexual behavior, and mental health outcomes, and used LCA to determine how many such risk profiles could be distinguished [@hopfer_social_2014].
Another explored heterogeneity in the population of military service members and veterans with prior suicide attempts,
and identified two latent profiles characterized by high versus low suicide risk based on self-reported suicidal ideation and information about prior attempted methods and severity
[@gromatsky_characteristics_2022].

@@

<!-- ### Controlling for covariates -->
<!-- CJ: I'm not sure we should call this a use case... in principle, -->
<!-- any time you're doing some kind of LCA you may want to predict group membership or predict something from group membership... -->
<!-- An extension of LCA is that containing covariates which can be used to -->
<!-- predict class membership. In this approach, we not only model the latent -->
<!-- class variable based on indicators, but we also relate the class -->
<!-- membership to other explanatory variables (@vermunt_latent_2017). An -->
<!-- example of using covariates comes from @nozadi_moderating_2016 who -->
<!-- applied LCA to identify the probability of children's membership to an -->
<!-- anxiety class. The authors tested several covariates including -->
<!-- children's age, sex, and accuracy scores. Age and sex were not found to -->
<!-- be related to the children's latent class membership, hence these -->
<!-- covariates were excluded from the analysis. Accuracy scores, however, -->
<!-- were related to probabilities of being in anxiety and attention-anxiety -->
<!-- classes. Therefore this covariate was kept as a valuable predictor. -->
<!-- Another example comes from @van_lissa_divergence_2015 who included sex  -->
<!-- as a predictor of the developmental trajectories and class membership in  -->
<!-- the previously discussed adolescent empathy development study. The rationale -->
<!-- was that adding sex as a covariate would account for known sex differences  -->
<!-- in empathy development, and lead to a model with lower bias. -->

<!-- ### Distal outcomes -->
<!-- CJ: I'm not sure we should call this a use case... in principle, -->
<!-- any time you're doing some kind of LCA you may want to predict group membership or predict something from group membership... -->
<!-- When our interest is the prediction of one or more outcomes, LCA can be used  -->
<!-- to construct latent classes as categorical predictors. @lanza_latent_2013  -->
<!-- demonstrated how LCA can be used to classify adolescents into depression  -->
<!-- classes, and subsequently these classes can be used to predict smoking, grades, -->
<!-- and delinquency. The study showed that the outcomes predicted by class  -->
<!-- membership can be binary (regular smoking), continuous (grades) or  -->
<!-- count (delinquency). In the previously discussed study of Belgian adolescent -->
<!-- identities, @maene_perceived_2022 first constructed the latent variable -->
<!-- representing different multiple identification strategies. Subsequently,  -->
<!-- the most likely class membership based on posterior class probability was -->
<!-- used as an "observed" categorical variable for prediction of depressive symptoms. -->

@@

The best practices in estimation, as outlined in Table
\@ref(tab:checkest), are rooted in existing recommendations for best
practices for estimating specific subtypes of LCA,
including latent
class growth analysis [@van_de_schoot_grolts-checklist_2017] and latent
class analysis with ordinal indicators [e.g., @nylund-gibson_ten_2018].
These were generalized to be relevant to all types of LCA, and updated
to current best practices, as explained below.
In order to aid researchers working with latent trajectory models,
@van_de_schoot_grolts-checklist_2017 developed a protocol called
Guidelines for Reporting on Latent Trajectory Studies (GRoLTS).

@@

@@

Examining observed data is essential for any analysis as it may reveal
patterns and violations of assumptions that had not been considered
prior to data collection. Special attention should be paid to level of
measurement of the indicators.
Finite Gaussian mixture models (including
LPA) are only suitable for continuous variables.
Indicators with an
ordinal level of measurement are likely to violate the assumption of
within-class normal distributions of mixture models [see
@vermunt_k-means_2011].
Personal experience consulting on LCA methods
and moderating the `tidyLPA` Google group suggest that the
misapplication of mixture models to ordinal (e.g., Likert-type)
indicators is the most common source of user error. Whereas it has been
argued that some parametric methods are robust when scales with 7+
response categories are treated as continuous [e.g., @norman_likert_2010],
it is very unlikely
that an ordinal variable can be modeled as a *mixture* of multiple
normal distributions.
To illustrate the problem, consider what happens when estimating a 5-class solution on a 5-point Likert scale.
Each class-specific mean could describe a single response
category,
and a class-specific variance component would be nonsensical.
In sum, Likert-type scales are rarely suitable for Gaussian mixture modeling;
latent class analysis with ordinal indicators is more appropriate.

@@

Previous work has emphasized the importance of examining the pattern of
missing data and reporting how missingness was handled
[@van_de_schoot_grolts-checklist_2017].
While we concur that investigating missingness is due diligence,
it is important to emphasize that missingness is adequately handled by default in `OpenMx`, which is the computational backend of `tidySEM`.
Its Full Information Maximum Likelihood (FIML) estimator makes use of all available
information without imputing missing values [@endersRelativePerformanceFull2001].
FIML is a best-practice solution for handling missing data;
on par with multiple imputation when the
estimation model includes the same variables as the imputation model
[@lee_comparison_2021].
Although FIML does not, by
default, handle missingness in exogenous variables - all indicator
variables in LCA are endogenous, so this is not a concern for LCA.
Multiple imputation, a common alternative to FIML,
is less suitable to LCA for two reasons. First,
because LCA methods are often computationally expensive,
and conducting them on multiple imputed datasets may be unfeasible.
Second, because many multiple imputation methods have some assumption of normality, which is likely to be violated when it is assumed that the data come from a heterogenous population.

Three types of missingness have
been distinguished in the literature [@rubin_inference_1976; 
@enders_applied_2022]: random missingness (MCAR);
missingness contingent on *observed* variables that can thus be known and controlled for (MAR);
and missingness related to unobserved variables, which can neither be known nor controlled for (MNAR).
A so-called "MCAR" test is often reported in relation to missingness
[@jamshidian_tests_2010].
But note that this name is somewhat misleading,
as a significant test indicates that missingness is MAR,
but a non-significant test statistic does not distinguish between MCAR or MNAR.
As the classic MCAR test relies on the comparison of variances across groups with different patterns of missing data, it assumes normality [@little_test_1988].
This assumption is tenuous in the context of LCA.
A non-parametric MCAR test may be more suitable [@jamshidian_tests_2010].
For this tutorial,
the lead author has implemented this test in the `mice` package as `mice::mcar()`, with a
fast backend in C++ and new printing and plotting methods.
Our recommendation is to inspect missingness and report the proportion of missingness per variable using `tidySEM::descriptives()`,
and testing for MAR using `mice::mcar()`,
before proceeding with FIML.

One minor concern is that the K-means algorithm, which
`tidySEM` uses to determine starting values, is *not* robust to
partially missing data [@hartiganAlgorithm136KMeans1979].
When missing data causes it to fail,
`tidySEM` automatically switches to hierarchical clustering [@murtaghWardHierarchicalAgglomerative2014].
The user can also specify a different
clustering algorithm or specify starting values manually.
Cases with complete missing data should be removed,
or their data should be imputed another way.

### Model Specification

Before estimating a LCA model, researchers should carefully specify the model,
based on theoretical and empirical considerations.
This will ensure that the model accurately represents the data and provides reliable results.
Two factors play a role in LCA model specification:
the number of latent classes and the parametrization within each class.
The total number of parameters of a latent class model consists of the number of parameters within each class times the number of classes,
plus a classification vector with $K-1$ free elements.
Different types of latent class models have different parameters.
For example, latent profile analyses typically assume class-specific means,
equal variances across classes, and zero covariances (i.e., indicators are independent after accounting for latent class membership), see `mx_profiles()`.
LCA with ordinal indicators typically assume free thresholds across classes,
where each threshold reflects a proportion of responses within a particular response category, see `mx_lca()`.
Like latent profile analyses, latent class growth analyses typically have free means across classes, but with respect to the latent growth variables instead of the observed variables, see `mx_growth_mixture()`.
In the general sense, any structural equation model can be specified,
either using `lavaan` model syntax or using `OpenMx` RAM specification,
and supplied to the general function `mx_mixture()`.
The resulting model may have means, (co)variances, regression coefficients, factor loadings, and thresholds for ordinal variables.
All of these parameters can be freely estimated,
constrained to be equal across classes using equality constraints (e.g., `x ~~ a*x` fixes the variance of `x` to be equal across classes),
or fixed to a certain value (e.g., `x ~.5*1` fixes the mean of `x` to `.5`).

As any of these parameters could be misspecified,
prior literature has emphasized the importance of considering alternative model specifications [@van_de_schoot_grolts-checklist_2017].
While we agree with this recommendation,
two considerations have remained underemphasized in prior literature.
One consideration is that alternative model specifications may be approached differently depending on whether the LCA is data driven (exploratory), or theory driven
(confirmatory).
The section on Use Cases illustrates these different applications.

Prior literature has focused on exploratory LCA.
In this context, model specification typically consists of an exhaustive grid search along varying numbers of classes and model specifications.
The "correct" model specification is then determined based on a combination of fit indices, significance tests, and interpretability [@nylund_deciding_2007].
Confirmatory LCA does not require such an extensive search over the model space.
Theoretical models can be compared to a smaller set.
If hypothesized models have been specified a priori,
and possibly even preregistered,
it makes sense to focus on the estimation and interpretation of these models, rather than proceed in a purely data-driven manner.
One key challenge is that LCA lacks objective fit indices,
and the fit of a theoretical model can thus only be determined relative to other likely candidates.
For this reason, the theoretical model should still be compared to a few others.
A one-class model is a sensible benchmark, to test whether LCA is even appropriate.
A theoretical model would be supported by a substantial (and significant) difference in fit with a one-class model.
Other candidates for comparison include competing theoretical models and those with a few more or less classes than the hypothesized number.
Here again, a substantial difference in fit supports the theoretical model - but note that small differences in fit do not give cause to reject a theoretical model.
The theoretical model may still be preferred, for example, to ensure comparability with prior work.
The fact that the theoretical model received little differential support should of course be discussed,
along with potential implications for future work.
The BIC is the preferred fit index for comparing

<!-- HIER -->
<!-- For latent profile analysis, the function -->
<!-- `tidySEM::mx_profiles(classes, variances, covariances)` largely -->
<!-- automates this process. The argument `classes` indicates which class -->
<!-- solutions should be estimated (e.g., 1 through 6). The argument -->
<!-- `variances` specifies whether variances should be `"equal"` or -->
<!-- `"varying"` across classes. The argument `covariances` specifies whether -->
<!-- covariances should be constrained to `"zero"`, `"equal"` or `"varying"` -->
<!-- across classes. The means are free to vary across classes by default, -->
<!-- although the more general function `tidySEM::mx_mixture()` could be used -->
<!-- to circumvent this. After all models have been estimated, the function -->
<!-- `tidySEM::table_fit()` can be used to obtain a model fit table suitable -->
<!-- for determining the optimal model according to best practices. Note, -->
<!-- however, that this table does not include the bootstrapped likelihood -->
<!-- ratio test (BLRT) by default, because this test is very computationally -->
<!-- expensive. It is recommended to use the function `tidySEM::BLRT()` to -->
<!-- compare a shortlist of likely candidate models based on other fit -->
<!-- indices. More on fit indices can be found in the Model Fit Indices -->
<!-- subsection of this paper. -->

The second consideration is the risk of overfitting.
We address each in turn.

As explained before, the total number of parameters scales with the number of estimated classes.
Consequently, LCA methods have a potentially very high number of parameters.

@@

Among studies using LCA, reporting practices vary significantly
[@weller_latent_2020]. Various authors have tried to improve and
standardize ways of reporting LCA [e.g. @masyn_latent_2013;
@weller_latent_2020], but more work is needed. 
<!-- This does not belong here: -->
<!-- @van_lissa_worcs_2020 -->
<!-- developed WORCS, a workflow for open reproducible code in science. WORCS -->
<!-- consists of step-by-step guidelines for research projects based on the -->
<!-- TOP-guidelines developed by @nosek_promoting_2015. WORCS workflow can be -->
<!-- easily implemented in R in form of an R package which facilitates -->
<!-- preregistration, article drafting, version control, citation and -->
<!-- formatting, among others [@van_lissa_worcs_2020]. -->

TOP-guidelines emphasize the use of comprehensive citation (including
referencing the software used in the analysis), as well as code and data
sharing wherever possible [@nosek_promoting_2015]. @van_lissa_worcs_2020
If the original data cannot be shared, sharing synthetic data is a good alternative.
The function `worcs::synthetic()` provides a generic, non-parametric method to create a synthetic copy of data.
When synthesizing data for LCA, it is also possible to synthesize data directly from the model.
LCA models constructed in `tidySEM` are estimated using `OpenMx`,
and as such, can benefit from all existing functions in that package.
To synthesize data from a mixture model `mix`, use the function `mxGenerateData(mix)`.

the entire research project is made reproducible so that others may
download it and reproduce it with just one click; for guidance, see
@van_lissa_worcs_2020.

@@

```{r, include = TRUE, eval=F}
# Load required packages
library(tidySEM) 
# Load data
df <- data_mix_ordinal
```

@@

```{r, include = TRUE, eval=F}
tidySEM::descriptives(df)
```

@@

Before we fit a series of LCA models, we set a random seed using
`set.seed()`. This is an important
step as there is some inherent randomness in the LCA computations, and
having the same seed number ensures that two separate researcher obtain
exactly the same results when fitting LCA models.

Finally, we reach the step of fitting LCA models.
As all variables are ordinal, we can use the convenience function
`tidySEM::mx_lca()`, which is a wrapper for the generic function `mx_mixture()`,
optimized for latent class analysis with ordinal data.
Any mixture model can be specified through `mx_mixture()`,
but at the time of writing, there are two more wrapper functions for specialized models:
`mx_profiles()`, for latent profile analysis, and `mx_growth_mixture()`, for latent growth analysis and growth mixture models.
All of these functions have arguments `data` and number of `classes`.
All variables in `data` are included in the analysis,
so relevant variables must be selected first.
Below, we fit 1 to 4 class solutions.
<!-- Depending on your computer's computational power, -->
<!-- this might take a while. -->
<!-- As `tidySEM::mx_lca()` run several models at a time, the resulting object `res`  -->
<!-- is a `mixture_list` class where each $K$ class solution is a list element, or -->
<!-- more specifically an `OpenMx` model (`MxModel`). -->

@@

<!-- Notice that the model run 10 times for each $K$ class solution. -->
<!-- This way we can choose the best fitting model across all the solutions. -->
<!-- In this example, we see that all models converged without issues.  -->
<!-- We can inspect the class of resulting objects. -->

```{r, include = TRUE, eval=F, echo = F}
is(res)
is(res[[1]])
```

@@

Description <- c(
    "Number of classes $K$",
    "Basic fit measure of the natural log of the likelihood of the data",
    "Total number of model parameters (in all classes)",
    "Relative fit measure: Akaike Information Criterion",
    "Relative fit measure: Bayesian Information Criterion",
    "Lowest average posterior class probability for class cases were assigned to",
    "Highest average posterior class probability for class cases were assigned to",
    "Proportion of cases in the smallest class based on posterior class probability",
    "Proportion of cases in the largest class based on posterior class probability")
